Bad forecast, good decision: predictive accuracy is not everything

### 2020-11-19 manuscript/manuscript.Rmd
edits

---
title: "Bad forecast, good decision: predictive accuracy is not everything"
date: "`r Sys.Date()`"
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
address:
  - code: ucb
    address: "ESPM Department, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA"
abstract: |
    With the rapid expansion of available data, it is frequently suggested that _forecasting performance_ is the ultimate test of a scientific model.  Yet the model that makes the most accurate forecasts is not always the model that yields the best decisions.  This is no accident of statistics, even formal criteria such as strictly proper scoring rules cannot determine the best model for decision-making because they do not consider the context of the decision process itself.  I illustrate this paradox using a classic example of ecological decision making in fisheries management.  This example also reveals how an understanding of the decision problem can both explain the poor management outcomes of model which produces the best forecast, and also see how to avoid making such a mistake.  As forecasting becomes an increasing attractive and viable possibility for assessing ecological models and applying them to real-world decisions, I hope this illustration serves as a reminder that modeling can not be done in a vacuum, but must reflect the context in which the model will be applied.
journal: TBD
layout: 3p
bibliography: refs.bib
csl: nature.csl
output: 
  rticles::elsevier_article:
    dev: cairo_pdf
    keep_tex: true
keywords: 
  - transients, 
  - optimal control, 
  - adaptive management, 
  - stochasticity, 
  - uncertainty, 
  - ecological management
preamble: |
  \journal{preprint}
  \usepackage{lineno}
  \linenumbers

@@

A primary purpose of statistical analyses and ecological modeling is to make forecasts for the future [@Clark2001; @Gneiting2014].  Accurate forecasts are important both in assessing the accuracy of our models and understanding of natural processes, they also underpin policy and decision making [@Dietze2018].  Yet the model that leads to the best decisions is not always the model that makes the most accurate forecasts, as I illustrate here.  Ecological processes are intrinsically complex, so that even our best models can only ever be approximations of underlying processes.  Even strictly proper scoring rules for probabilistic forecasts [@Gneiting2007] will not always select the best model to guide decision-making.  Surprisingly, this can even happen when a decision is derived directly from a complex optimization routine of a probabilistic predictive models [@Marescot2013].  Here, I use a classic, well-understood example from fisheries management [@Schaefer1954; @Clark1973; @Reed1979] to illustrate both the paradox of how a model with the worst forecast provides the best decision outcomes, as well as show how we can avoid selecting models that are poorly suited for management by considering the management context more explicitly. These results underscore that in choosing the best model for decision-making, it can be more important to capture a single key feature of the process than it is to make the most accurate prediction about future states.

@@

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from model 1 result in a significantly higher stock size than under model 2 (panel A).  Despite this less agressive harvest policy, economic returns under model 1 are substantially higher (panel B)."}
fig2a + fig2b
```

@@

```{r figure3, fig.width=7, fig.height=4, fig.cap="Panel A: Population recruitment curves for each model, compared to that of the true model. Panel B: The computed optimal policy of each model, derived by SDP, expressed in terms of the target escapement (population size remaining after harvest) for each possible stock size. Model 1 overharvests consistently, while the target escapement under model 2 is nearly identical to that of model 3."}
fig3a + ggtitle("A") + fig3b + ggtitle("B")
```

>>>>

---
title: "Bad forecast, good decision: predictive accuracy is not everything"
date: "`r Sys.Date()`"
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
address:
  - code: ucb
    address: "ESPM Department, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA"
abstract: |
     Model-based forecasts have been enabled by recent explosion of data and computational methods and spurred on by decision-makers appetite for forecasts in everything from elections to pandemic response. It is taken for granted that the model which makes the most accurate forecast, accounting for uncertainty, will also be the best model to inform decision-making.  Using a classic example from fisheries management, I demonstrate that selecting the model that produces the best forecast can lead to decidedly worse outcomes. This situation can arise whenever the models are only approximations of the reality, underscoring the risk of evaluating models only by prediction accuracy and not decision outcomes
journal: " "
layout: 3p
bibliography: refs.bib
csl: science.csl
output: word_document
#  rticles::elsevier_article:
#    dev: cairo_pdf
#    keep_tex: true

keywords: 
  - transients, 
  - optimal control, 
  - adaptive management, 
  - stochasticity, 
  - uncertainty, 
  - ecological management
preamble: |
  \journal{preprint}
  \usepackage{lineno}
  \linenumbers

@@

A primary purpose of statistical analyses and ecological modeling is to make forecasts for the future [@Clark2001; @Gneiting2014].  Accurate forecasts are important both in assessing the accuracy of our models and understanding of natural processes, they also underpin policy and decision making [@Dietze2018]. Strictly proper scoring rules provide mathematical guarantees that they will, on average, select the true model over any other[@Gneiting2007], paving the way for ever more complex and more accurate forecasting models. Yet the model that leads to the best decisions is not always the model that makes the most accurate forecasts, as I illustrate here. Surprisingly, this can even happen when a decision is derived directly from a complex optimization routine of a probabilistic predictive models [@Marescot2013].  Reality is complex; even our best models can only ever be approximations of underlying processes.  Here, I use a classic, well-understood example from fisheries management [@Schaefer1954; @Clark1973; @Reed1979] to illustrate both the paradox of how a model with the worst forecast provides the best decision outcomes, as well as show how we can avoid selecting models that are poorly suited for management by considering the management context more explicitly. These results underscore that in choosing the best model for decision-making, it can be more important to capture a single key feature of the process than it is to make the most accurate prediction about future states.

@@

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from Model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under Model 1 are also substantially higher (panel B)"}
fig2a + fig2b
```

@@

```{r figure3, fig.width=7, fig.height=4, fig.cap="Panel A: Population recruitment curves for each model, compared to that of the true Model. 3 Model 2 more closely approximates the true Model 3, but note that maximum value Model 1 and Model 3 occur at nearly the same value for the state, x.  Panel B: The computed optimal policy of each model, derived by SDP, expressed in terms of the target escapement (population size remaining after harvest) for each possible stock size. Model 1 over-harvests consistently, while the target escapement under Model 2 is nearly identical to that of the true Model 3."}
fig3a + ggtitle("A") + fig3b + ggtitle("B")
```
